**RMarkdown Project 2**
============================================

This is my markdown file that records all steps that I did for the second project of the Reproducible Research course on Coursera 

## 0. Download and read the data
```{r, echo = TRUE}
fileurl <- "C:/Users/ASUS/Documents/RepData_PeerAssessment1/repdata_data_activity/activity.csv"
dt <- read.csv(fileurl, header = TRUE, sep = ",")
#Change the format of date
dt$date <- as.Date(dt$date)
#Remove the NA values
dt <- subset(dt, !is.na(steps), 1:3)
```

## 1. Make a histogram of the total number of steps taken each day
```{r, echo = TRUE}
total.steps.per.day <- aggregate(dt$steps, by = list(dt$date), FUN = sum)
names(total.steps.per.day) <- c("Date", "Total.steps")
head(total.steps.per.day)
hist(total.steps.per.day$Total.steps, 
     xlab = "Number of total steps/day",
     main = "Histogram of the total number of steps taken each day")
```

## 2. Calculate and report the mean and median total number of steps taken per day
```{r, echo = TRUE}
Steps.report <- data.frame(Mean = mean(total.steps.per.day$Total.steps), 
                           Median = median(total.steps.per.day$Total.steps))
Steps.report
```

## 3. Make a time series plot
```{r, echo = TRUE}
avg.steps.per.interval <- aggregate(dt$steps, by = list(dt$interval), FUN = mean)
names(avg.steps.per.interval) <- c("Interval", "Average.step")

with(avg.steps.per.interval, {
     plot(Interval, Average.step, 
     type = "l",
     xlab = "Interval",
     ylab = "Average number of steps",
     main = "Time series plot of the 5-minute interval and the average number of steps taken")
     })
```


##. 4. Which 5-minute interval contains the maximum number of steps?
```{r, echo = TRUE}
max.step <- max(avg.steps.per.interval$Average.step)
avg.steps.per.interval$Interval[which(avg.steps.per.interval$Average.step == max.step)]
```

##. 5. Calculate and report the total number of missing values in the dataset
``` {r, echo = TRUE}
dt.full <- read.csv(fileurl, header = TRUE, sep = ",")
dt.full$date <- as.Date(dt.full$date)
sum(is.na(dt.full$steps))
```

##. 6. Devise a strategy for filling in all of the missing values in the dataset
I will use the mean for that 5-minute interval to fill in the missing values

##. 7. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo = TRUE}
for (i in 1:length(dt.full$interval)) {
     if (is.na(dt.full$steps[i]) == TRUE) {
          j <- which(avg.steps.per.interval$Interval == dt.full$interval[i])
          dt.full$steps[i] <- avg.steps.per.interval$Average.step[j]
     }
     else {
          next
     }
}
```

##. 8. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r, echo = TRUE}
total.steps.per.day.filled <- aggregate(dt.full$steps, by = list(dt.full$date), FUN = sum)
names(total.steps.per.day.filled) <- c("Date", "Total.steps")
head(total.steps.per.day.filled)
hist(total.steps.per.day.filled$Total.steps, 
     xlab = "Number of total steps/day",
     main = "Histogram of the total number of steps taken each day")

Steps.report.filled <- data.frame(Mean = mean(total.steps.per.day.filled$Total.steps), 
                           Median = median(total.steps.per.day.filled$Total.steps))
Steps.report.filled  
```

##. 9. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day
```{r, echo = TRUE}
w <- c()
weekend <- c("Saturday", "Sunday")
for (i in 1:length(dt.full$date)) {
     if (weekdays.Date(dt.full$date[i]) %in% weekend) {
          w <- c(w, "weekend")
     }
     else {
          w <- c(w, "weekday")
     }
}
dt.full$datetype <- w
dt.full$datetype <- as.factor(dt.full$datetype)
```

##. 10. Make a panel plot containing a time series plot 
```{r, echo = TRUE}
avg.steps.interval.w <- aggregate(steps ~ interval + datetype, data = dt.full, mean)
library(lattice)
xyplot(steps ~ interval | datetype, data = avg.steps.interval.w, 
      type = "l", 
      layout = c(1, 2),
      xlab = "5-minute interval", 
      ylab = "Average number of steps",
      main = "Average Number of Steps Taken (across all weekday days or weekend days)")
```

